Supplemental Material for ``Assessing United States county-level exposure for research on tropical cyclones and human health"

### 2020-09-09 ehp_revision/response_to_reviewers.Rmd

\begin{quotation}
\noindent
\textit{"The rainfall data are generally well-correlated with ground-based
observations, but may sometimes underestimate very high rainfall values (Figure
2)."}
\end{quotation}
\end{shaded}

**Response:**

This is a great suggestion. In the original manuscript, we describe this lower
correlation at higher precipitation values through the use of examples from
specific storms, and discuss the potential reasons for it. Measuring the
correlation specifically within these higher-precipitation events would add more
quantitative evidence to this discussion.

We have added estimates of these correlations specifically for
high-precipitation events in a new table in the Supplemental Material (Table S2
of the manuscript, reproduced as Table \@ref(tab:highprecipcorr) of this
response document).

@@

\input{tables/tornado_misclass.tex}

We have added four new tables in the Supplemental Material, one for each of the
hazard-based metrics we considered. We have reproduced these tables in this
response document as Tables \@ref(tab:misclasswind)--\@ref(tab:misclassflood)
Each breaks down agreement and disagreement between the distance-based proxy
metric and that hazard-based metric within several storms in our study period.
These tables focus on storms with at least 250 counties exposed by at least one
of the metrics considered. These tables allow readers both to investigate
general patterns in disagreement and exposure misclassification across these
storms and also to see cases of unusual storms. For example, for most storms,
disagreement between the distance proxy metric and the wind-based metric
resulted mainly from mis-classification of counties as exposed based on distance
while they did not experience gale-force winds. However, there were a few
storms, like Hurricane Bertha in 1996, where the inverse was the case, and a
distance-based metric would have missed a large number of counties that
experienced gale-force winds from the storm (misclassified them as unexposed),
but not misclassified many counties as exposed when they in fact did not
experience winds at this level.

We have added a reference to these new tables in the main text of the manuscript:

> **Text in the revised manuscript (relevant addition in bold):** 
"We drew similar conclusions when we investigated all 46 tropical cyclones
between 1996 and 2011 (when data for all five metrics were available) for which
100 or more counties were exposed based on at least one metric (Figure 7;
**Tables S3--S6 provide underlying numbers for the most extensive of
these, the storms for which 250 of more counties were exposed based on at least
one metric)**."

> **Text in the revised manuscript (relevant addition in bold):**
"Based on our results, the use of a distance-based metric to assess exposure to
any of these hazards, or the use of measurements from one hazard as a proxy for
exposure to any of the other hazards considered, would often introduce exposure
misclassification **(Tables S3--S6)**."

@@

For all outcomes, there are various reasonable choices that can be picked for
the threshold in defining exposure. There is not room to go into these details
within Table 1, so instead we added a separate table to the Supplemental
Material that provides both some reasoning behind the thresholds selected in
this study and also other reasonable thresholds that could be considered.
This is Table S1 of the revised manuscript and has been reproduced as 
Table \@ref(tab:thresholds) at the end of this response document.

\input{tables/thresholds_table.tex}

Ultimately, this choice will depend on what a researcher considers as plausible
causal pathways for the outcome he or she is studying. For example, if power
outages form a plausible step in the pathway, then the threshold for gale-force
or strong gale-force winds might be reasonable, while if the pathway is more
likely to be through larger-scale building destruction, a higher threshold would
be more reasonable. Some studies have done sensitivity analyses to consider how
results changed using different exposure metrics. For example, @czajkowski2011
performed a sensitivity analysis using different assumptions about wind decay
away from the center of a storm, while @currie2013, who used a distance-based
exposure metric, considered different buffer distances as a sensitivity
analysis.

We have added text to introduce this supplemental table in the revised manuscript, 
both in the Methods and in the Discussion:

> **Text in the revised Methods (added text in bold)**:
"For other exposure metrics, each county was
classified as exposed to a tropical cyclone based on whether the exposure
metric exceeded a certain threshold (Table~\ref{tab:exposuremetrics}). We
picked reasonable thresholds (e.g., the threshold for gale-force wind for wind
exposure; **Table S1**), but others could be used with the open-source data and its
associated software."

> **Text in the revised Discussion (added text in bold)**:
"Finally, to assess patterns and agreement for binary exposure classifications,
we have chosen one set of sensible thresholds for binary classifications based
on continuous metrics (rainfall, maximum sustained surface wind, and distance
from the storm's track), **but other thresholds would be reasonable
depending on hypothesized pathways for a given epidemiological study (Table
S1)**."

@@

\medskip

\begin{shaded}
\textbf{EHP Request 5:}
Figures 5--6: Please check to see if the colors used can be distinguished if
printed in black and white, and if not, use a “color-blind friendly” palette if
possible. If this is not feasible, we can request and exemption from
accessibility requirements, but we try to do this only when alternatives  would
degrade the information content of  the figure.
\end{shaded}

**Response:**

As we note in our response to EHP Request 2, we used colormaps for all figures
that are designed to be perceptible both for those with common types of color
vision deficiency and when printed in grayscale Figures \@ref(fig:fig5check1)
and \@ref(fig:fig5check2) at the end of this response show what Figure 5 of the
main text would look like under three common types of color vision deficiency
and when printed in black and white. Figures \@ref(fig:fig6check1) and
\@ref(fig:fig6check2) at the end of this response show the same thing for Figure
6 of the main text.

For Figure 6 of the main text, for some types of color balance, it could be
unclear which of the two color legends pairs with which part of the graph. We
have therefore added a note to the figure legend to clarify:

> **Figure legend for Figure 6 in revised manuscript (added text shown in bold)**: 
"Agreement between exposure classifications based on different single-hazard
exposure metrics for all storms between~1996 and~2011 for which at least~100
counties were exposed based on at least one metric. Each row shows one storm,
and the color of each cell shows the measured Jaccard index for each pair of
exposure metrics (proportion of counties classified as exposed by both metrics
out of counties classified as exposed by either metric). For Grace in 2003 and
Ida in 2009, there were no county exposures for either the tornado-based metric
or the wind-based metric (indicated by gray squares). Colors to the right of the
heatmap show the number of exposed counties based on any of the metrics, **and
this panel is linked with the color scale labeled '# of counties exposed by
any metric'.** Storms are displayed within clusters that have similar patterns
based on hierarchical clustering."

These figures should therefore not require any exemption from accessibility
requirements, as they should satisfy these requirements.

### 2020-09-09 ehp_revision/writing/ehp_discussion.tex
Work some more on revisions

Based on our results, the use of a distance-based metric to assess exposure to
any of these hazards, or the use of measurements from one hazard as a proxy for
exposure to any of the other hazards considered, would often introduce exposure
misclassification \textbf{(Tables S[x]--S[x])}. This conclusion reinforces
similar findings from a study of Florida's 2004 storm
season~\parencite{grabich2015measuring}.  For some studies, such exposure
misclassification might plausibly be differential.  For example, tropical
cyclone wind exposures tend to be concentrated in counties near the coast,
while tropical cyclone rain exposures sometimes extend well inland.  If the
etiologically-relevant exposure for a health outcome is extreme rainfall but
exposure is classified based on measurements of wind, the probability of being
misclassified as unexposed would be higher in inland counties, while the
probability of being misclassified as exposed would be higher in coastal
counties. If coastal counties differ from inland counties in either the outcome
of interest or in factors associated with risk of that outcome, exposure
misclassification would be differential~\parencite{savitz2016interpreting}.
Such differential exposure misclassification could bias estimates of tropical
cyclone effects either towards the null (estimating a lower or null association
compared to the true association that exists) or away from the null (estimating
a larger association than actually exists)~\parencite{savitz2016interpreting,
armstrong1998effect}.

@@

Finally, to assess patterns and agreement for binary exposure classifications,
we have chosen one set of sensible thresholds for binary classifications based
on continuous metrics (rainfall, maximum sustained surface wind, and distance
from the storm's track).  Results and conclusions would differ somewhat with
other threshold choices. We have published code for this analysis online
(\url{https://github.com/geanders/county_hurricane_exposure}), allowing other
researchers to explore other threshold choices for these analyses.

>>>>

Based on our results, the use of a distance-based metric to assess exposure to
any of these hazards, or the use of measurements from one hazard as a proxy for
exposure to any of the other hazards considered, would often introduce exposure
misclassification \textbf{(Tables S3--S6)}. This conclusion reinforces
similar findings from a study of Florida's 2004 storm
season~\parencite{grabich2015measuring}.  For some studies, such exposure
misclassification might plausibly be differential.  For example, tropical
cyclone wind exposures tend to be concentrated in counties near the coast,
while tropical cyclone rain exposures sometimes extend well inland.  If the
etiologically-relevant exposure for a health outcome is extreme rainfall but
exposure is classified based on measurements of wind, the probability of being
misclassified as unexposed would be higher in inland counties, while the
probability of being misclassified as exposed would be higher in coastal
counties. If coastal counties differ from inland counties in either the outcome
of interest or in factors associated with risk of that outcome, exposure
misclassification would be differential~\parencite{savitz2016interpreting}.
Such differential exposure misclassification could bias estimates of tropical
cyclone effects either towards the null (estimating a lower or null association
compared to the true association that exists) or away from the null (estimating
a larger association than actually exists)~\parencite{savitz2016interpreting,
armstrong1998effect}.

@@

Finally, to assess patterns and agreement for binary exposure classifications,
we have chosen one set of sensible thresholds for binary classifications based
on continuous metrics (rainfall, maximum sustained surface wind, and distance
from the storm's track), \textbf{but other thresholds would be reasonable
depending on hypothesized pathways for a given epidemiological study (Table
S1)}.  Results and conclusions would differ somewhat with other threshold
choices. We have published code for this analysis online
(\url{https://github.com/geanders/county_hurricane_exposure}), allowing other
researchers to explore other threshold choices for these analyses.

### 2020-09-09 ehp_revision/writing/ehp_methods.tex
Work some more on revisions

For our study domain, we used counties in the eastern half of the \ac{US}
(Figure~\ref{fig:hurrtracks}).  We included all tropical cyclones between 1988
and 2018 that were tracked in \ac{HURDAT2} \parencite{landsea2013} and came
within~250 \si{\kilo\metre} of at least one eastern \ac{US} county
(Figure~\ref{fig:hurrtracks}). We therefore including all land-falling or
near-land Atlantic basin tropical cyclones while excluding storms that never
neared the \ac{US}.  \subsection*{Distance-based exposure metric}

We first measured how close each storm's central track came to each county. We
used tracking data from \ac{HURDAT2}, which records the storm center's position
at four standardized times for weather data collection (synoptic times),
6:00~am, 12:00~pm, 6:00~pm, and 12:00~am \ac{UTC}.  We interpolated this
position to~15-\si{\minute} intervals using natural cubic spline
interpolation~\parencite{hurricaneexposure}. At each~15-\si{\minute} interval,
we measured the distance between the storm's center and each county's
population mean center, **as of the 2010 US Decennial Census**~\parencite{countycenters}. We took the minimum distance
for each county as a measure of how close the storm came to the county over its
lifetime.  We also recorded the time when this occurred for each county so we
could later link observed data on precipitation, flooding, and tornadoes.  To
allow matching with data recorded in local time (e.g., health data), we
converted these times from \ac{UTC} to local time~\parencite{countytimezones}.

@@

Two of the exposure metrics (flood- and tornado-based) were inherently binary
in our data, since these metrics were based on whether an event was listed in
the NOAA Storm Events database.  For other exposure metrics, each county was
classified as exposed to a tropical cyclone based on whether the exposure
metric exceeded a certain threshold (Table~\ref{tab:exposuremetrics}). We
picked reasonable thresholds (e.g., the threshold for gale-force wind for 
wind exposure), but others could be used with the open-source data and its
associated software.

>>>>

For our study domain, we used counties in the eastern half of the \ac{US}
(Figure~\ref{fig:hurrtracks}).  We included all tropical cyclones between 1988
and 2018 that were tracked in \ac{HURDAT2} \parencite{landsea2013} and came
within~250 \si{\kilo\metre} of at least one eastern \ac{US} county
(Figure~\ref{fig:hurrtracks}). We therefore including all land-falling or
near-land Atlantic basin tropical cyclones while excluding storms that never
neared the \ac{US}.

\subsection*{Distance-based exposure metric}

We first measured how close each storm's central track came to each county. We
used tracking data from \ac{HURDAT2}, which records the storm center's position
at four standardized times for weather data collection (synoptic times),
6:00~am, 12:00~pm, 6:00~pm, and 12:00~am \ac{UTC}.  We interpolated this
position to~15-\si{\minute} intervals using natural cubic spline
interpolation~\parencite{hurricaneexposure}. At each~15-\si{\minute} interval,
we measured the distance between the storm's center and each county's
population mean center, \textbf{as of the 2010 US Decennial
Census}~\parencite{countycenters}. We took the minimum distance for each county
as a measure of how close the storm came to the county over its lifetime.  We
also recorded the time when this occurred for each county so we could later
link observed data on precipitation, flooding, and tornadoes.  To allow
matching with data recorded in local time (e.g., health data), we converted
these times from \ac{UTC} to local time~\parencite{countytimezones}.

@@

Two of the exposure metrics (flood- and tornado-based) were inherently binary
in our data, since these metrics were based on whether an event was listed in
the NOAA Storm Events database.  For other exposure metrics, each county was
classified as exposed to a tropical cyclone based on whether the exposure
metric exceeded a certain threshold (Table~\ref{tab:exposuremetrics}). We
picked reasonable thresholds (e.g., the threshold for gale-force wind for wind
exposure; \textbf{Table S1}), but others could be used with the open-source data and its
associated software.

### 2020-09-09 ehp_revision/writing/ehp_results.tex
Work some more on revisions

We drew similar conclusions when we investigated all~46 tropical cyclones
between~1996 and~2011 (when data for all five metrics were available) for
which~100 or more counties were exposed based on at least one metric
(Figure~\ref{fig:jaccard}). In this figure, each row provides results for one
tropical cyclone, and each box in that row shows the Jaccard coefficient for a
pair of metrics. For all pairs of metrics, agreement in exposure assessment
was, at best, moderate for all but a few storms. When comparing distance- and
wind-based exposure assessment, only about 10\% of storms had Jaccard indices
higher than 0.6 (i.e., out of the counties assessed as exposed by at least one
of the two metrics in the pair,~60\si{\percent} or more were assessed the same
under both metrics). For comparisons of assessments based on other combinations
of distance-, wind-, rain-, and flood-based metrics, fewer than 5\% of storms
had Jaccard indices above 0.6.  The tornado-based metric had universally poor
agreement with other metrics in county-level classification across the tropical
cyclones considered.

>>>>

We drew similar conclusions when we investigated all~46 tropical cyclones
between~1996 and~2011 (when data for all five metrics were available) for
which~100 or more counties were exposed based on at least one metric
(Figure~\ref{fig:jaccard}; \textbf{Tables S3--S6 provide underlying
numbers for the most extensive of these, the storms for which 250 of more
counties were exposed based on at least one metric}). In this figure, each row
provides results for one tropical cyclone, and each box in that row shows the
Jaccard coefficient for a pair of metrics. For all pairs of metrics, agreement
in exposure assessment was, at best, moderate for all but a few storms. When
comparing distance- and wind-based exposure assessment, only about 10\% of
storms had Jaccard indices higher than 0.6 (i.e., out of the counties assessed
as exposed by at least one of the two metrics in the pair,~60\si{\percent} or
more were assessed the same under both metrics). For comparisons of assessments
based on other combinations of distance-, wind-, rain-, and flood-based
metrics, fewer than 5\% of storms had Jaccard indices above 0.6.  The
tornado-based metric had universally poor agreement with other metrics in
county-level classification across the tropical cyclones considered.
